Reduced plasticity and variance in physiological rates for terrestrial ectotherm populations under climate change

### 2024-01-22 ms.qmd
All Alex's comments dealt with

Monthly average temperature data were extracted from the ERA5 climate model, available from the Copernicus climate data store [@Hersbach2020-hx]. For each population and species in the dataset we extracted a 30-year period (1950-2022) of either surface air temperature at 2 meters (0.01 resolution) for both terrestrial and freshwater taxa, or sea surface temperature for the marine taxa (at 0.25 resolution) using the *ncdf4* R package [vers. `r utils::packageVersion("ncdf4")`, @Pierce2021-lc]. We chose a 2-meter resolution because we believed that it more likely reflects the micro-thermal environment experienced by terrestrial and freshwater ectotherms at those locations. For terrestrial species we also collected soil temperature as for many species this maybe more representative of microhabitat choice compared to air temperature. We fit models using both air and soil temperature and found that the results were qualitatively similar. As such, we only present results for air temperature.

Using the thermal time-series data for each location we summarised various metrics of thermal variability across months and years as well as estimates of thermal predictability (i.e., autocorrelation). To estimate thermal variability, we calculated the coefficient of variation ($\frac{SD}{M}$, where SD = standard deviation in temperature and M = the mean temperature for each year). To estimate thermal predictability, we calculated the auto-regressive time lag across the entire dataset. Theoretical and empirical studies of plasticity evolution have emphasised the importance of both climate variability and predictability in plasticity evolution.

We also extracted climate projections into the future to understand the effects that climate warming will have on physiological rate variance. We used the CanESM2 climate model (2005-2100) under a high emissions scenerio (RCP8.5).

@@

```{r, fig-1}
#| label: fig-1
#| fig-cap: "Taxonomic distribution of acute and acclimation $Q_{10}$ estimates across major habitats. **A)** Phylogenetic distribution of taxa contained within the data. The total number of acute and acclimation type Q10 effect sizes are highlighted as well as whether the taxa is marine, freshwater or terrestrial. Silouettes are representative taxa of major clades within the tree. **B)** Acute and Acclimation $lnRR_{Q_{10}}$ across marine, freshwater, and terrestrial environments. **C)** $lnCVR_{Q_{10}}$ across traits for marine, freshwater and terrestrial systems. Note there were no differences between acute and acclimation $Q_{10}$ types. k = total number of effect size estimates while the numbers in brackets indicate the number of species."
#| warning: false
#| echo: false

@@

Thermal variability (i.e., $CV$) and predictability a population experienced in its past did not explain acclimation capacity or changes in physiological rate variance among terrestrial, marine or freshwater populations (@fig-4; @fig-5).

>>>>

Monthly average temperature data were extracted from the ERA5 climate model, available from the Copernicus climate data store [@Hersbach2020-hx]. For each population and species in the dataset we extracted a 30-year period (1950-2022) of either surface air temperature (0.01$^{\circ}$ resolution) for both terrestrial and freshwater taxa, or sea surface temperature for the marine taxa (at 0.25$^{\circ}$ resolution) using the *ncdf4* R package [vers. `r utils::packageVersion("ncdf4")`, @Pierce2021-lc]. We chose surface temperature because we believed that it was more likely reflects the micro-thermal environment experienced by terrestrial and freshwater ectotherms at those locations. For terrestrial species we estimated soil temperatures as this maybe more representative of microhabitat choice compared to air temperature. We fit models using both air and soil temperature and found that the results were qualitatively similar. As such, we only present results for air temperature.

Using the thermal time-series data for each location we calculated metrics of thermal variability across months and years as well as estimates of thermal predictability (i.e., autocorrelation). To estimate thermal variability, we calculated the coefficient of variation ($\frac{SD}{M}$, where SD = standard deviation in temperature and M = the mean temperature for each year). To estimate thermal predictability, we calculated the auto-regressive time lag across the entire dataset. Theoretical and empirical studies of plasticity evolution have emphasised the importance of both climate variability and predictability in plasticity evolution.

Lastly, to illustrate the effects that climate warming could have on physiological rate variance we also extracted climate projections and calculated thermal variability and predictability for the future. We used the CanESM2 climate model (2005-2100) under a high emissions scenerio (RCP8.5).

@@

```{r, fig-1}
#| label: fig-1
#| fig-cap: "Taxonomic distribution of acute and acclimation $Q_{10}$ estimates across major habitats. **A)** Phylogenetic distribution of taxa contained within the data. The total number of acute and acclimation type Q10 effect sizes are highlighted as well as whether the taxa is marine, freshwater or terrestrial. Silouettes are representative taxa of major clades within the tree. **B)** Acute and Acclimation $lnRR_{Q_{10}}$ across marine, freshwater, and terrestrial systems. **C)** $lnCVR_{Q_{10}}$ across traits for marine, freshwater and terrestrial systems. Note there were no differences between acute and acclimation $Q_{10}$ types. k = total number of effect size estimates while the numbers in brackets indicate the number of species."
#| warning: false
#| echo: false

@@

Thermal variability (i.e., $CV$) and predictability experienced by a population in the past did not explain acclimation capacity or changes in physiological rate variance among terrestrial, marine or freshwater populations (@fig-4; @fig-5).
